---
title: "GLM"
author: "Robert Bartolini"
date: "12/26/2020"
output: 
  html_document: 
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

https://github.com/rjmaitri/GLM.git

## 1. Comparing Means
To start with, let’s warm up with a simple one-way ANOVA model. This example, from Whitlock and Schluter chapter 15 question 22 looks at the mass of lodgepole pinecones from different habitats.

1.1. Load and plot the data. Choose a plot that not only shows the raw data, but also the means and SE or CI of those means. +1 EC if Michael thinks it’s fancy.
```{r}
###Load the data and Look into the df
LP_pine <- read.csv("https://www.zoology.ubc.ca/~whitlock/ABD/teaching/datasets/15/15q22LodgepolePineCones.csv")

reactable::reactable(LP_pine)

```

```{r}
###Create a boxplot that shows the mean and +/- SE, raw data

#Function to calculate +/- SE from the mean *credit:https://stackoverflow.com/questions/25999677/how-to-plot-mean-and-standard-error-in-boxplot-in-r
MinMeanSEMMax <- function(x) {
  v <- c(min(x), mean(x) - sd(x)/sqrt(length(x)), mean(x), mean(x) + sd(x)/sqrt(length(x)), max(x))
  names(v) <- c("ymin", "lower", "middle", "upper", "ymax")
  v
}

#####################################################################
#boxplot the stats function and add dotplot for raw data

p <- ggplot(LP_pine, aes(factor(habitat), conemass, color = habitat))+
  stat_summary(fun.data=MinMeanSEMMax, geom="boxplot")+ 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1)+
  ggtitle("Pinecone Mass By Habitat Stats: Min, Mean +/- 1SE, Max and Raw Data")+
  xlab("Habitat")
  
p



```

1.2 Fit a model using least squares and evaluate all relevant assumptions. List them out as you test them. Can we use this model? If not, fix it. But if we can, no fix is needed!

```{r}
#create a factor variable for habitat
# creating the factor variable
LP_pine$habitat.f <- factor(LP_pine$habitat)
#Verify it worked
is.factor(LP_pine$habitat.f)

```

```{r}
#Check results
LP_pine$habitat.f
```

Excellent!

```{r}
#build a linear model using Least Squares Regression
LS_pine <- lm(formula = conemass ~ habitat.f, data = LP_pine)

#Lets see what the fit looks like:
summary(LS_pine)

```

1.2 How much variation is explained by your model?

<p style="color:green">The multiple and adjusted R2 values are nearly 0.86-0.88, meaning that the island habitat is highly explanatory of conemass.</span>



1.3 Show which means are different from each other. Are you correcting p-values? If so, how, and justify your choice.

```{r}


```